### 2018-04-13 paper/approach.tex
added less formal explanations to intro and approach

\subsection{Clause-Dependence Disjoint Systems}
\label{sec:CDD-defn}
%
The key contribution of our work is the introduction of the class of
Clause-Dependence Disjoint (CDD) systems.
%
CDD systems can be solved
in co-NP time and contains the union of all previous introduced classes
that can be solved in co-NP time.
%
\begin{defn}
  \label{defn:cdds}
  For a given recursion-free CHC system $\cc{S}$,
  %
  if for all sibling pairs, $P, Q \in \predof{S}$,
  %
  the transitive dependencies of $P$ and $Q$ are disjoint ($\tdepsof{P} \cap
  \tdepsof{Q} = \emptyset$)
  %
  and no predicate shows more than once in the body of a single clause,
  %
  then $\cc{S}$ is \emph{Clause-Dependence Disjoint (CDD)}.
\end{defn}
%
CDD systems model hierarchical programs with branches and procedure calls such
that each execution path invokes each statement at most once.
%
\begin{ex}
  The CHC system $\mcchc$ is a CDD system. An argument is given in
  \autoref{sec:solve-ex}.
\end{ex}

@@

%
\subsubsection{Constructing constraints for predicates}
%
For a CDD system $S$, a predicate $P \in \predof{S}$, and a partial
solution $\sigma$ that maps predicates to their solutions, the formula
$\ctrof{P,\sigma}$ is a compact representation of the constraints of
$P$.
%
If $\sigma$ does not contain $P$, then the constraint of $P$ is
constructed from the clauses where $P$ is the head.  When $\sigma$
does contain $P$, $\ctrof{P,\sigma}$ is a lookup from $\sigma$:
%
\[
  \ctrof{P,\sigma}=
  \begin{dcases}
    \biglor_{(\cc{C}_i \in \cc{S}) \land (\headof{C_i} = P)}
    \left( \consof{C_i} \land %
      \bigland_{ \cc{Q} \in \bodyof{C_i}} \cc{b}_Q
    \right),
  &\text{if } P \notin \sigma\\
  \sigma[P], &\text{if } P \in \sigma
  \end{dcases}
\]
%
% introduce the isused predicates:
Each $P \in \predof{S}$ has a corresponding boolean variable $\cc{b}_P$.
%
The counterexample characterization of $P$ constrains that if $P$ is used
(meaning $\cc{b}_{P} = True$), then its constraint must hold:
\[
  \vc{P,\sigma} = \neg \cc{b}_{P} \lor \ctrof{P,\sigma}
\]

\begin{ex}
  \label{ex:ctr}
  When $\solvecdd$ solves predicate $\cc{L}_9$ in $\mcchc$, it generates a
  constraint based on clauses (5) and (6):
  $$\ctrof{\cc{L}_9,\sigma} =
    (\cc{abs'} = \cc{n} \land \cc{b}_{\cc{L}_6})
    \lor
    (\cc{abs'} = -\cc{n} \land \cc{b}_{\cc{L}_8})$$
  The counterexample characterization for $\cc{L}_9$ is based on its boolean
  indicator and this constraint:
  $$\vc{\cc{L}_9,\sigma} = \neg \cc{b}_{\cc{L}_9} \lor \ctrof{\cc{L}_9,\sigma}$$
\end{ex}

@@

% core algorithm
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{For a recursion-free CHC system $S$.}
  % Output: inductive invariants.
  \Output{A solution to $S$ or $\none$.}
  % expand procedure:
  \myproc{$\sys(S)$ %
    \label{line:shara-begin}}{ %
    % construct expansion of C:
    $(S', \eta) \assign %
    \expand(S)$ \label{line:shara-expand} \; %
    % expand, solve, case split:
    \Switch{ $\solvecdd(S' )$ %
      \label{line:shara-case} }{
      % case: result is no solution:
      \lCase{ $\none$: }{ \Return{$\none$} \label{line:shara-ret-none} } %
      % case: result is a solution:
      \lCase{ $\sigma'$: }{ %
        \Return{ $\collapse{ \eta }{ \sigma' }$ } %
        \label{line:shara-ret-collapse}
      } %
    } %
  } %
  %
  \caption{\sys: a solver for recursion-free CHCs, which uses
    procedures $\expand{}$ (see \autoref{app:cons-cdd}) and
    $\solvecdd$ (see
    \autoref{sec:solve-cdd}). }
  \label{alg:shara}
\end{algorithm}
% walk through the sys algorithm:
\sys (\autoref{alg:shara}), given a recursion-free CHC system $S$
(\autoref{line:shara-begin}), 
returns a solution
to $S$ or the value $\none$ to denote that $S$ is
unsolvable.
%
\sys first runs a procedure $\expand$ on $S$ to obtain a CDD expansion
$S'$ of $S$.
%
\sys then invokes $\solvecdd$ on $S'$.
%
When \solvecdd~returns that $S'$ has
no solution, \sys~propogates $\none$ (\autoref{line:shara-ret-none}).
%
Otherwise, \sys constructs a solution from the CDD solution, $\sigma'$, by
invoking $\collapse{ \eta }{ \sigma' }$ (\autoref{line:shara-ret-collapse}).

% define collapse:
For $\sigma'$ and %
correspondence $\eta$ from $P' \in \predof{S'}$ to $P \in \predof{S}$,
$\collapse{\eta}{\sigma'}$ maps solution $\sigma'$ to
$\sigma$[$P \to \bigland_{ \eta(P') = P}
\sigma'(P')$].
%
$\expand$ is given in \autoref{app:cons-cdd}.

>>>>

\subsection{Clause-Dependence Disjoint Systems}
\label{sec:CDD-defn}
%
The key contribution of our work is the introduction of the class of
Clause-Dependence Disjoint (CDD) CHC systems:
%
\begin{defn}
  \label{defn:cdds}
  For a given recursion-free CHC system $\cc{S}$,
  %
  if for all sibling pairs, $P, Q \in \predof{S}$,
  %
  the transitive dependencies of $P$ and $Q$ are disjoint ($\tdepsof{P} \cap
  \tdepsof{Q} = \emptyset$)
  %
  and no predicate shows more than once in the body of a single clause,
  %
  then $\cc{S}$ is \emph{Clause-Dependence Disjoint (CDD)}.
\end{defn}
%
CDD systems model hierarchical programs with branches and procedure calls such
that each execution path invokes each statement at most once.
%
\begin{ex}
  The CHC system $\mcchc$ is a CDD system. An argument is given in
  \autoref{sec:solve-ex}.
\end{ex}

@@

%
\subsubsection{Constructing constraints for predicates}
%
In order to construct efficiently sized pre and post formulas for
relational predicates, we need a method of compactly expressing the
constraints on a given predicate.
%
For a CDD system $S$, a predicate $P \in \predof{S}$, and a partial
solution $\sigma$ that maps predicates to their solutions, the formula
$\ctrof{P,\sigma}$ is a compact representation of the constraints of
$P$.
%
If $\sigma$ does not contain $P$, then the constraint of $P$ is
constructed from the clauses where $P$ is the head.  When $\sigma$
does contain $P$, $\ctrof{P,\sigma}$ is a lookup from $\sigma$.
%
Each $P \in \predof{S}$ has a corresponding boolean variable
$\cc{b}_P$:
%
\[
  \ctrof{P,\sigma}=
  \begin{dcases}
    \biglor_{(\cc{C}_i \in \cc{S}) \land (\headof{C_i} = P)}
    \left( \consof{C_i} \land %
      \bigland_{ \cc{Q} \in \bodyof{C_i}} \cc{b}_Q
    \right),
  &\text{if } P \notin \sigma\\
  \sigma[P], &\text{if } P \in \sigma
  \end{dcases}
\]
%
% introduce the isused predicates:
The counterexample characterization of $P$ is a small extension of the
compact constraint of $P$.
%
It states that if $P$ is used
(meaning $\cc{b}_{P} = True$), then the constraint of $P$ must hold:
\[
  \vc{P,\sigma} = \neg \cc{b}_{P} \lor \ctrof{P,\sigma}
\]

\begin{ex}
  \label{ex:ctr}
  When $\solvecdd$ solves predicate $\cc{L}_9$ in $\mcchc$, it generates a
  constraint based on clauses (5) and (6):
  $$\ctrof{\cc{L}_9,\sigma} =
    (\cc{abs'} = \cc{n} \land \cc{b}_{\cc{L}_6})
    \lor
    (\cc{abs'} = -\cc{n} \land \cc{b}_{\cc{L}_8})$$
  The counterexample characterization for $\cc{L}_9$ is based on its boolean
  indicator and its constraint:
  $$\vc{\cc{L}_9,\sigma} = \neg \cc{b}_{\cc{L}_9} \lor \ctrof{\cc{L}_9,\sigma}$$
\end{ex}

@@

% core algorithm
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{For a recursion-free CHC system $S$.}
  % Output: inductive invariants.
  \Output{A solution to $S$ or $\none$.}
  % expand procedure:
  \myproc{$\sys(S)$ %
    \label{line:shara-begin}}{ %
    % construct expansion of C:
    $(S', \eta) \assign %
    \expand(S)$ \label{line:shara-expand} \; %
    % expand, solve, case split:
    \Switch{ $\solvecdd(S' )$ %
      \label{line:shara-case} }{
      % case: result is no solution:
      \lCase{ $\none$: }{ \Return{$\none$} \label{line:shara-ret-none} } %
      % case: result is a solution:
      \lCase{ $\sigma'$: }{ %
        \Return{ $\collapse{ \eta }{ \sigma' }$ } %
        \label{line:shara-ret-collapse}
      } %
    } %
  } %
  %
  \caption{\sys: a solver for recursion-free CHCs, which uses
    procedures $\expand{}$ (see \autoref{app:cons-cdd}) and
    $\solvecdd$ (see
    \autoref{sec:solve-cdd}). }
  \label{alg:shara}
\end{algorithm}
% walk through the sys algorithm:
\sys (\autoref{alg:shara}), given a recursion-free CHC system $S$
(\autoref{line:shara-begin}), 
returns a solution
to $S$ or the value $\none$ to denote that $S$ is
unsolvable.
%
\sys first runs a procedure $\expand$ on $S$ to obtain a CDD expansion
$S'$ of $S$.
%
\sys then invokes $\solvecdd$ on $S'$.
%
When \solvecdd~returns that $S'$ has
no solution, \sys~propogates $\none$ (\autoref{line:shara-ret-none}).
%
Otherwise, \sys constructs a solution from the CDD solution,
$\sigma'$, by invoking $\collapse{ \eta }{ \sigma' }$
(\autoref{line:shara-ret-collapse}).
%
$\textsc{Collapse}$ is designed to convert the solution for the CDD system
back to a sensible solution for the original problem. It does this by
taking the conjunction of all interpretations of predicates which
correspond to one predicate in the original problem.
%
That is, given a CDD solution $\sigma'$ and a correspondence $\eta$ from
$P' \in \predof{S'}$ to $P \in \predof{S}$,
$\collapse{\eta}{\sigma'}$ generates an entry in $\sigma$ for each
predicate in the original system:
$\sigma[P] \assign \bigland_{ \eta(P') = P} \sigma'(P')$.
%
$\expand$ is given in \autoref{app:cons-cdd}.

### 2018-04-13 paper/introduction.tex
added less formal explanations to intro and approach

% intro:
\section{Introduction}
\label{sec:intro}
% context, problem defn:
Many critical problems in program verification can be reduced to
solving systems of Constrained Horn Clauses (CHCs), a class of
logic-programming
problems~\cite{bjorner13,flanagan03,rummer13a,rummer13b}.
%
A CHC is a logical implication where the antecedent is a conjunction
of uninterpreted predicates with a fact in a background
theory and the consequent is an uninterpreted predicate.
%
In the context of a CHC we call the antecedent the body and the
consequent the head.  A CHC system is a set of CHCs. 
%
The goal of a CHC solving problem is to find suitable interpretations
for each predicate such that each CHC is logically
consistent in isolation.

% introduce the problem of solving recursion-free systems:
In this work we focus on the subclass of CHC systems which are known
as \emph{recursion-free}. In a recursion-free CHC system, the
derivation of each predicate will not contain
itself.
%
Recursion-free CHC systems are an important subclass for two reasons.
%
First, recursion-free systems can be used to model safety properties
for hierarchical programs~\cite{lal-qadeer15,lal-qadeer-lahiri12}
(programs with only bounded iteration and recursion).
%
Second and most importantly, a well-known approach for solving a
general CHC system reduces the input problem to solving a sequence
of recursion-free systems.
%
Such approaches attempt to synthesize a solution for the   original
system from the solutions of recursion-free systems~\cite{bjorner13}.
%
The performance of such solvers relies
heavily on the performance of solving recursion-free CHC systems.
%

>>>>

% intro:
\section{Introduction}
\label{sec:intro}
% context, problem defn:
Many critical problems in program verification can be reduced to
solving systems of Constrained Horn Clauses (CHCs), a class of
logic-programming
problems~\cite{bjorner13,flanagan03,rummer13a,rummer13b}.
%
A CHC is a logical implication with the following form:
$$
  R_1(\vec{v_1}) \leftarrow R_2(\vec{v_2}) \land R_3(\vec{v_3}) \land
  ... \land \varphi(\vec{v_0}, \vec{v_1}, \vec{v_2}, \vec{v_3},...)
$$
Here, the left side of the implication, called the head, contains an
uninterpreted relational predicate applied to a vector of variables.
%
The right side has any number of such predicates conjoined together
with a \emph{fact} ($\varphi$). The fact is a logical formula in a background
theory and may use variables named by the predicates.
%
The goal of a CHC solving problem is to find suitable interpretations
for each predicate such that each CHC is logically
consistent in isolation.

% introduce the problem of solving recursion-free systems:
In this work we focus on the subclass of CHC systems which are known
as \emph{recursion-free}. In a recursion-free CHC system, the
derivation of each predicate will not contain itself.
%
Less formally, a recursion-free CHC system is one where following
implication arrows through the system will never reach the same clause
twice.
%
Recursion-free CHC systems are an important subclass for two reasons.
%
First, recursion-free systems can be used to model safety properties
for hierarchical programs~\cite{lal-qadeer15,lal-qadeer-lahiri12}
(programs with only bounded iteration and recursion).
%
Second and most importantly, a well-known approach for solving a
general CHC system reduces the input problem to solving a sequence
of recursion-free systems.
%
Such approaches attempt to synthesize a solution for the   original
system from the solutions of recursion-free systems~\cite{bjorner13}.
%
The performance of such solvers relies
heavily on the performance of solving recursion-free CHC systems.
%
